---
title: "Google Case Study A"
author: "Goder"
date: "2025-01-16"
output: html_document
---

# Setting Up Environment
Setting up environment by load the package "tidyverse" and "conflicted" to manage conflict

```{r}
# Load tidyverse for data manipulation and visualization
library(tidyverse)

# Load conflicted to resolve naming conflicts between packages
library(conflicted)

# set dplyr::filter and lag as the default choice
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```

# Collect Data
Dataset from [divy_data]<https://divvy-tripdata.s3.amazonaws.com/index.html>. In this case, dataset used is from 2024, January to December

```{r}
# Load monthly Divvy trip data CSV files into separate data frames
DT_01 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202401-divvy-tripdata.csv")
DT_02 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202402-divvy-tripdata.csv")
DT_03 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202403-divvy-tripdata.csv")
DT_04 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202404-divvy-tripdata.csv")
DT_05 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202405-divvy-tripdata.csv")
DT_06 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202406-divvy-tripdata.csv")
DT_07 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202407-divvy-tripdata.csv")
DT_08 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202408-divvy-tripdata.csv")
DT_09 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202409-divvy-tripdata.csv")
DT_10 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202410-divvy-tripdata.csv")
DT_11 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202411-divvy-tripdata.csv")
DT_12 <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/202412-divvy-tripdata.csv")
```

## Check Structure Dataset
```{r}
str(DT_01)
str(DT_02)
```

## Merged all dataset into one big dataframe
```{r}
all_trips <- bind_rows(DT_01,DT_02,DT_03,DT_04,DT_05,DT_06,DT_07,DT_08,DT_09,DT_10,DT_11,DT_12)
```

## Remove start_station_id, end_station_name, end_station_id, start_lat, start_lng, end_lat, end_lng
```{r}
fix_all_trips <- all_trips %>% 
  select(-c(start_station_id, end_station_name, end_station_id, start_lat, start_lng, end_lat, end_lng))
```

```{r}
str(fix_all_trips) #See list of columns and data types (numeric, character, etc)
```

# Clean Data
This step describes cleaning up irrelevant data and adding new columns for further analysis

## Inspect Dataframe
```{r}
str(fix_all_trips)  #See list of columns and data types (numeric, character, etc)
head(fix_all_trips)  #See the first 6 rows of data frame
```

## Check for empty values (NA) in each column
```{r}
colSums(is.na(fix_all_trips))
```

## Check unique value column member_casual
```{r}
length(unique(fix_all_trips$member_casual))
```
## Create new column named "day_of_week" that contains the days based on the date and column "month"
```{r}
# Convert the 'started_at' column to Date format
date_format <- as.Date(fix_all_trips$started_at)

# Extract the month name from the date and store it in a new column 'month'
fix_all_trips$month <- format(as.Date(date_format), "%B")

# Extract the day of the week from the date and store it in a new column 'day_of_week'
fix_all_trips$day_of_week <- format(as.Date(date_format), "%A")
```

```{r}
head(fix_all_trips$month) #See the first 6 rows of month column
head(fix_all_trips$day_of_week) #See the first 6 rows of day_of_week column
```

## Create new column named "ride_length" that contains value calculation of the length of time riding.

```{r}
# Calculate the difference in time between 'ended_at' and 'started_at' columns
difftime_ride <- difftime(fix_all_trips$ended_at, fix_all_trips$started_at)

# Convert the time difference to numeric format
difftime_ride <- as.numeric(as.character(difftime_ride))
```

```{r}
# Add the calculated ride length to the dataset as a new column 'ride_length'
fix_all_trips$ride_length <- difftime_ride

# Check if the 'ride_length' column is numeric
is.numeric(fix_all_trips$ride_length)
```


```{r}
# Check if any ride length in the 'ride_length' column is less than 0
any(fix_all_trips$ride_length < 0)
```

```{r}
# Check for any duplicated row names in the 'fix_all_trips' dataset
anyDuplicated(rownames(fix_all_trips))
```


## Save clean data for further analysis

```{r}
# Filter out rows with negative ride lengths and create a new dataset 'all_trips_2024'
all_trips_2024 <- fix_all_trips[!(fix_all_trips$ride_length < 0), ]
```


```{r}
str(all_trips_2024) #See list of columns and data types (numeric, character, etc)
```

```{r}
# Save the 'all_trips_2024' dataset to a CSV file without row names
write.csv(all_trips_2024, file = '2024_trips.csv', row.names = FALSE)
```

# Conduct Descriptive Analysis

```{r}
# Read the '2024_trips.csv' file into the 'all_trips' dataset
all_trips <- read.csv("D:/Berkas/Coursera Certification/Google Data Abalytics Professional Certification/Study Case/Case A/Dataset/2024_trips.csv")
```

```{r}
str(all_trips) #See list of columns and data types (numeric, character, etc)  
head(all_trips) #See the first 6 rows of data frame
```

## Total member and casual 
```{r}
# Group the data by 'member_casual' and summarize the number of rides for each group
all_trips %>% group_by(member_casual) %>% summarise(number_of_ride = n())
```


## Descriptive analysis on month (total rider type)

```{r}
# Convert 'month' column to a factor with levels in chronological order
all_trips$month <- factor(all_trips$month, 
                          levels = c("January", "February", "March", "April", "May",
                                     "June", "July", "August", "September", "October",
                                     "November", "December"))
```


```{r}
# Group the data by month and summarize the number of rides, then arrange by month
all_trips %>% group_by(month) %>% summarise(number_of_rides = n()) %>% arrange(month)

# Aggregate the data by month and rider type to count the number of ride IDs, then arrange by month
all_trips %>% aggregate(ride_id ~ month + member_casual, FUN = length) %>% arrange(month)
```

## Descriptive analysis on day_of_week

```{r}
# Convert 'day_of_week' column to a factor with levels in the correct order (Sunday to Saturday)
all_trips$day_of_week <- factor(all_trips$day_of_week, 
                          levels = c("Sunday", "Monday", "Tuesday", "Wednesday",
                                     "Thursday", "Friday", "Saturday"))
```


```{r}
# Group the data by day of the week, summarize the number of rides, and arrange by day of the week
all_trips %>% group_by(day_of_week) %>% summarise(number_of_rides = n()) %>% arrange(day_of_week)

# Aggregate the data by day of the week and rider type to count the number of ride IDs, then arrange by day of the week
all_trips %>% aggregate(ride_id ~ day_of_week + member_casual, FUN = length) %>% arrange(day_of_week)
```


## Descriptive analysis on ride_length (all figures in seconds)

```{r}
# Generate summary statistics for the 'ride_length' column in the 'all_trips' dataset
summary(all_trips$ride_length)
```
```{r}
# Compare members and casual users
# Calculate the mean ride length for each rider type
all_trips %>% aggregate(ride_length ~ member_casual, FUN = mean)

# Calculate the median ride length for each rider type
all_trips %>% aggregate(ride_length ~ member_casual, FUN = median)

# Calculate the maximum ride length for each rider type
all_trips %>% aggregate(ride_length ~ member_casual, FUN = max)

# Calculate the minimum ride length for each rider type
all_trips %>% aggregate(ride_length ~ member_casual, FUN = min)

# Count the number of rides for each rider type
all_trips %>% aggregate(ride_length ~ member_casual, FUN = length)
```

```{r}
# Calculate the average ride length grouped by rider type and day of the week, then arrange by member type and day of the week
all_trips %>% aggregate(ride_length ~ member_casual + day_of_week, FUN = mean) %>% arrange(member_casual, day_of_week)
```

## Descriptive Analysis on start_station_name

```{r}
# Count the number of unique start station names in the 'all_trips' dataset
length(unique(all_trips$start_station_name))

# Check if there are any empty start station names
any(all_trips$start_station_name == "")
```

```{r}
# Group the data by 'start_station_name', count the number of rides for each station, 
# arrange by the most visited, and select the top 10 most visited stations
all_trips %>% group_by(start_station_name) %>% summarise(most_visited = n()) %>% 
  arrange(desc(most_visited)) %>% slice(1:10)
```

# Visualization Result

```{r}
# Group data by month and membership type, count the number of rides, and arrange by month and membership type
all_trips %>%
  group_by(month, member_casual) %>%
  summarise(number_of_rides = n()) %>%
  arrange(month, member_casual) %>%

  # Create a bar plot showing the number of rides per month for each membership type
  ggplot(mapping = aes(x = month, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge") +  # Dodge position to separate bars for each membership type
  labs(title = "Number of Rides per Month",  # Plot title
       x = "Month",  # X-axis label
       y = "Number of Rides",  # Y-axis label
       fill = "Rider Type") +  # Legend title
  theme_minimal()  # Apply a minimal theme for the plot
```


```{r}
# Create a bar plot showing the number of rides per day of the week, separated by rider type
ggplot(data = all_trips, mapping = aes(x = day_of_week)) +
  geom_bar(aes(fill = member_casual), position = "dodge") +  # Separate bars by 'member_casual' for each day
  labs(title = "Number of Rides per Day",  # Plot title
       x = "Day",  # X-axis label
       y = "Number of Rides",  # Y-axis label
       fill = "Rider Type") +  # Legend title for 'fill' (rider type)
  theme_minimal()  # Apply a minimal theme for the plot
```
```{r message=FALSE, warning=FALSE}
# Group the data by membership type and day of the week, then calculate the number of rides and average duration
all_trips %>% 
  group_by(member_casual, day_of_week) %>% 
  summarise(number_of_rides = n(),  # Count the number of rides
            average_duration = mean(ride_length)) %>%  # Calculate the average ride duration
  arrange(member_casual, day_of_week) %>%  # Sort by membership type and day of the week
  
  # Create a bar plot of average duration per day, grouped by rider type
  ggplot(aes(x = day_of_week, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge") +  # Separate bars for each rider type
  labs(title = "Average Duration per Day",  # Plot title
       x = "Day",  # X-axis label
       y = "Average Duration",  # Y-axis label
       fill = "Rider Type") +  # Legend title for fill (rider type)
  theme_minimal()  # Apply a minimal theme for the plot
```

```{r}
# Group the data by rider type, then calculate the number of rides and average ride duration for each type
all_trips %>% 
  group_by(member_casual) %>% 
  summarise(number_of_rides = n(),  # Count the number of rides for each rider type
            average_duration = mean(ride_length)) %>%  # Calculate average ride duration
  arrange(member_casual) %>%  # Arrange by rider type
  
  # Create a bar plot for average duration per rider type
  ggplot(aes(x = member_casual, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge") +  # Create a dodge bar chart
  labs(title = "Average Duration per Rider Type",  # Plot title
       x = "Rider Type",  # X-axis label
       y = "Average Duration",  # Y-axis label (corrected from 'Number of Rides')
       fill = "Rider Type") +  # Legend title for rider type
  theme_minimal()  # Apply a minimal theme for a clean plot appearance
```

```{r}
# Group the data by start station name, count the number of visits per station, and arrange by most visited
all_trips %>% 
  group_by(start_station_name) %>% 
  summarise(most_visited = n()) %>%  # Count visits per station
  arrange(desc(most_visited)) %>%  # Arrange stations by most visited
  slice(2:6) %>%  # Select the 2nd to 6th most visited stations
  
  # Create a bar plot of the most visited stations
  ggplot(aes(x = start_station_name, y = most_visited, fill = start_station_name)) +
  geom_col(position = "dodge") +  # Create dodge bar chart
  labs(title = "Most Visited Station Name",  # Plot title
       x = "Station Name",  # X-axis label
       y = "Number of Visits",  # Y-axis label
       fill = "Station Name") +  # Legend title for station name
  theme_minimal()  # Apply a minimal theme for the plot
```

## Export Summary For Further Analysis

Create a csv file that will be visualize in Tableau
```{r}
# Group the data by rider type, month, and day of the week, and calculate summary statistics
summarise_value <- all_trips %>% 
  group_by(member_casual, month, day_of_week) %>%  # Group by rider type, month, and day of the week
  summarise(
    number_of_rides = n(),  # Count the number of rides
    avarage_duration = mean(ride_length)  # Calculate the average ride duration
  ) %>% 
  group_by(member_casual, month, day_of_week)  # Reapply the grouping for further analysis if needed
```
```{r}
write.csv(summarise_value, "summarise_trips_2024.csv")
```

